### 2018-04-10 paper/approach.tex

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) systems.
%
Solving an CHC system in CDD is coNP time.
%
CDD strictly contains
the union of classes of recursion-free systems that can be solved
in coNP time that have been introduced in previous work.
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all pair of uninterpreted predicates $P, Q \in \predof{S}$ that are siblings(\autoref{sec:chcs}), the 
  transitive dependences of $P$ and $Q$ are disjoint $\tdepsof{P} \cap \tdepsof{Q} = \emptyset$, and there is 
  no uninterpreted predicate shows more than once in a body of clause, then $\cc{S}$ is
  \emph{Clause-Dependence-Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branching and
procedure calls that in each execution path that each statments be
exected at most once.
%
\QZ{rework example}
\begin{ex}
  The CHC system $\mcchc$ (given in \autoref{sec:solve-ex}) is a CDD
  system.
  %
  The only clauses in $\mcchc$ that contain multiple uninterpreted
  predicates are \autoref{eqn:then} and \autoref{eqn:else}, which both
  contain $\cc{L4}$ and $\cc{dbl}$.
  %
  The dependences of $\cc{L4}$ and $\cc{dbl}$ are disjoint.
\end{ex}

% talk about body disjoint
The class of CDD systems contains body-disjoint and linear
systems introduced in previous work.
%
For recursion-free system $\cc{S}$, if for each uninterpreted predicate
$Q \in \predof{S}$, there is at most one caluse $\cc{C}$ that $Q \in \bodyof{C}$,
and there is no uninterpreted predicate shows more than once in a body of clause, then
$\cc{S}$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
%
If the body of each clause in $\cc{S}$ contains at most
relational predicate, then $\cc{S}$ is
\emph{linear}~\cite{albarghouthi12a}.

@@

% introduce solveCDD
the procedure $\solvecdd$ (\autoref{alg:solve-cdd}) is a solver for CDD systems.
%
$\solvecdd$, given a CDD system $\cc{S}$, it first do a topological sort on all
uninterpreted predicates in $\cc{S}$ based on their dependency relations.
%
In the order of topological sort, for each uninterpreted predicates $\cc{P}$ 
 $\solvecdd$ calles $\solveitp$ to compute the binary interpolation of 
 encoded pre and post formulas with the current solution.
 %
 If the conjunction of pre and post formulas is satisifiable, then $\solvecdd$
 returns $\nosoln$ to indicate this CDD system $\cc{S}$ is not solvable.
 %
 Otherwise, $\solvecdd$ binds the uninterpreted predicate $\cc{P}$ to the
 interpolation result $\cc{I}$ as part of the soltuion.
 %
 The encoding procedure $\prectr$ and $\postctr$ are defined below.
 %
 $\solvecdd$ returns the soltuion $\sigma$.

\QZ{rework example}
%
\begin{ex}
  $\solvecdd{ \mcpreds }$, given $\mcchc$, may generate
  interpretations of $\mcpreds$ in any topological ordering of the
  dependency relation of $\mcchc$, depending on the relational
  predicates that it chooses at \autoref{line:case-deps}.
  %
  One such ordering is \cc{L3}, \cc{L4}, \cc{dbl}, \cc{L9}, \cc{main},
  $\mathsf{Err}$.
\end{ex}

\subsubsection{Constructing constrains of uninterpreted 
predicate}
%
For an uninterpreted predicate $\cc{P} \in \predof{S}$ and $\cc{S}$ is an CDD system  
,the formula $\ctrof{Q}$ is a compact representation of constrains of $\cc{P}$, defined as follows.
% introduce the isused predicates:
For each $\cc{Q} \in \predof{S}$, there is a corresponding boolean variable
$b_Q$.
% for each relational predicate, define constraint:
The constraint of $P$, denoted $\ctrof{P}$, constrains
that if $P$ occurs in a derivation of $S$, then there is
some clauses $C \in S$ in which $P$ is the head of $\cc{C}$
\textbf{(1)} each $Q \in \bodyof{C}$ is used and 
\textbf{(2)} the constrained of the clause $\consof{C}$ is encoded.
%
I.e., $\ctrof{P}$ is
\[\lnot b_Q \land 
\left(
\biglor_{ (\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
\left( \consof{C_i} \land %
\bigland_{ \cc{Q} \in \bodyof{C_i}} %
b_Q \right) \right)
\]
%
The counterexample characterization of $\mathcal{S'}$ constrains that
\textbf{(1)} the query predicate of $\mathcal{S'}$ must be used in a
derivation and %
\textbf{(2)} for $R \in \mathcal{R}$, if $R$ is used, then its
constraint must hold.
%
I.e., $\vc{ \mathcal{R} }(\mathcal{S'})$ is $\used{
  \queryof{\mathcal{S'} } } \land %
\bigland_{R \in \mathcal{R} } \left( \used{R} \implies \ctrof{R}
\right)$.

\QZ{rework example}
\begin{ex}
  % 
  $\vc{ \mcpreds }(\mcchc)$ characterizes all runs of \cc{multiCall}
  that result in an error.
  %
  Although the relational predicates \cc{L4} and \cc{dbl} occur in the
  bodies of multiple clauses, they are each modeled with only a single
  instance of logical variables.
  % 
  Such a collection is sufficient because each run of \cc{multiCall}
  reaches \cc{L4} and \cc{dbl} at most once.
  %
  Correspondingly, each derivation of $\mcchc$ contains at most one
  occurrence of relational predicates \cc{L4} and \cc{dbl}.
\end{ex}

% pre-constraint:
\subsubsection{Constructing the pre-constraint of a relational predicate}
\label{sec:cons-pre}
%
For a CDD system $\cc{S}$, given a uninterpreted predicate $\cc{P} \in \predof{\cc{S}}$
, because $\solvecdd$ solve each uninterpreted predicate in a toplogical order, 
it has a partial solution $\sigma : \cc{Q} \to \formula$ for all uninterpreted predicates in
$\depsof{P}$.
%
the pre-constraint $\prectr(\cc{P},\sigma)$ is a formula that the conjunction of the constrains
of $\cc{P}$ and solutions of all dependencies of $P$ with their boolean varialbe.
%
In particular, $\prectr(\cc{P},\sigma)$ is:
\[\ctrof{P} \land
\left(
\bigland_{\cc{Q} \in \depsof{P}}
\left(b_Q \land \sigma[ \cc{Q} ]  \right)
\right)
\]
%
\QZ{rework example}
\begin{ex}
  \label{ex:pre-ctr}
  Assume that $\mcsolve$, given $\mcchc$, synthesizes a partial
  solution $\sigma \in \interps{ \mcpreds }$ that interprets \cc{L3}
  as $\true$, and chooses to then find an interpretation of \cc{L4}.
  %
  The pre-constraint of \cc{L4} is $\true \land \cc{res}_{\cc{4}} =
  0$, where $\cc{res}_{\cc{4}}$ is a logical variable that models the
  value of \cc{res} at line \cc{4}.
\end{ex}

\subsubsection{Constructing the post-constraint of a relational
  predicate}
\label{sec:cons-ctx}
%
For $\sigma \in \interps{ \mathcal{R} }$ a partial solution of
$\mathcal{S}$ and $R \in \mathcal{R}$, the post-constraint $\postctr{
  \mathcal{R} }(\mathcal{S}, \sigma, R) \in \formulas$ is mutually
inconsistent with the interpretation of $R$ in each solution of
$\mathcal{S}$ that extends $\sigma$.
% dependents of R:
In particular, let $\mathcal{D}_0$ be the reflexive dependents of $R$
in $\mathcal{S}$,
% clause siblings of dependents:
let $\mathcal{D}_1$ be the siblings in $\mathcal{S}$ of
$\mathcal{D}_0$,
%
let $\mathcal{D}_2$ be all dependences of $\mathcal{D}_1$, and
% collect all predicates that we define a constraint over:
let $\mathcal{D} = \mathcal{D}_0 \union \mathcal{D}_1 \union
\mathcal{D}_2$.
% define constraint for each predicate
Let $\restrict{ \mathcal{S} }{ \mathcal{D} }$ be $\mathcal{S}$
restricted to clauses whose head's relational predicate is in
$\mathcal{D}$.
% define instantiation of C
The post constraint for $\mathcal{S}$ under $\sigma$ at $R$ is the
counterexample characterization of $\restrict{ \mathcal{S} }{
  \mathcal{D} }$ grounded on $\sigma$ (defined in \autoref{sec:chcs}).
%
I.e., $\postctr{ \mathcal{R} }(\mathcal{S}, \sigma, R) = \vc{
  \mathcal{R} }(\sigma( \restrict{ \mathcal{S} }{ \mathcal{D} }))$.
% example post constraint:
\begin{ex}
  \label{ex:ctx-ctr}
  Consider the case in which $\mcsolve$, given $\mcchc$, generates an
  interpretation of \cc{L4}, introduced in \autoref{ex:pre-ctr}.
  %
  $\mcsolve$ collects relational predicates $\mathcal{D}_0 = \elts{
    \cc{L4}, \cc{L9}, \cc{main}, \mathsf{Err} }$, $\mathcal{D}_1 =
  \elts{ \cc{dbl} }$, and $\mathcal{D}_2 = \emptyset$.
  %
  $\mcsolve$ constructs $\postctr{ \mcpreds }(\mcchc, \sigma,
  \cc{L4})$ to be the counterexample characterization of $\mcchc$
  restricted to clauses whose head relational predicate is in
  $\mathcal{D} = \elts{ \cc{L4}, \cc{L9}, \cc{main}, \mathsf{Err} }$.
\end{ex}

>>>>

\subsection{Clause-Dependence Disjoint Systems}
\label{sec:CDD-defn}
%
The key contribution of our work is the introduction of the class of
Clause-Dependence Disjoint (CDD) systems.
%
Solving an CHC system in CDD is coNP time.
%
CDD strictly contains
the union of classes of recursion-free systems that can be solved
in coNP time that have been introduced in previous work.
%
\begin{defn}
  \label{defn:cdds}
  For a given recursion-free CHC system $\cc{S}$,
  %
  if for all pair of uninterpreted predicates $P, Q \in \predof{S}$ that are siblings(\autoref{sec:chcs}), the 
  transitive dependences of $P$ and $Q$ are disjoint $\tdepsof{P} \cap \tdepsof{Q} = \emptyset$, and there is 
  no uninterpreted predicate shows more than once in a body of clause, then $\cc{S}$ is
  \emph{Clause-Dependence-Disjoint (CDD)}.
\end{defn}
%
CDD systems model hierarchical programs with branching and
procedure calls that in each execution path that each statments be
exected at most once.
%
\begin{ex}
  The CHC system $\mcchc$ (given in \autoref{sec:solve-ex}) is a CDD
  system.
  %
  The only clauses in $\mcchc$ that its body has multiple uninterpreted
  predicates are clause (7) that contains $\cc{L9}$ and $\cc{db}$.
  %
  The transitive dependences of $\cc{L9}$ and $\cc{db}$ are \{\cc{L6}\,\cc{L8}\,\cc{L4}\} and $\varnothing$, which are disjoint.
\end{ex}

% talk about body disjoint
The class of CDD systems contains tree structure system, and body-disjoint and linear
systems introduced in previous work.
%
Tree structure system is contained by body-disjoint system. 
%
For recursion-free system $S$, if for each uninterpreted predicate
$Q \in \predof{S}$, there is at most one caluse $C$ that $Q \in \bodyof{C}$,
and there is no uninterpreted predicate shows more than once in a body of clause, then
$S$ is \emph{body-disjoint}~\cite{rummer13a,rummer13b}.
%
If the body of each clause in $\cc{S}$ contains at most
relational predicate, then $\cc{S}$ is
\emph{linear}~\cite{albarghouthi12a}.

@@

% introduce solveCDD
the procedure $\solvecdd$ (\autoref{alg:solve-cdd}) is a solver for CDD systems.
%

$\solvecdd$, given a CDD system $\cc{S}$, it first do a topological sort on all
uninterpreted predicates in $\cc{S}$ based on their dependency relations.
%
In the order of topological sort, for each uninterpreted predicates $\cc{P}$ 
 $\solvecdd$ calles $\solveitp$ to compute the binary interpolation of 
 encoded pre and post formulas with the current solution.
 %
 If the conjunction of pre and post formulas is satisifiable, then $\solvecdd$
 returns $\nosoln$ to indicate this CDD system $\cc{S}$ is not solvable.
 %
 Otherwise, $\solvecdd$ binds the uninterpreted predicate $\cc{P}$ to the
 interpolation result $\cc{I}$ as part of the soltuion.
 %
 The encoding procedure $\prectr$ and $\postctr$ are defined below.
 %
 $\solvecdd$ returns the soltuion $\sigma$.

%
\begin{ex}
  $\solvecdd$, given $\mcchc$, may generate
  interpretations of $\mcpreds$ in any topological ordering of the
  dependency relation of $\mcchc$.
  %
  One such ordering is \cc{L4}, \cc{L6}, \cc{L8}, 
   \cc{L9}, \cc{db} and \cc{main}.
\end{ex}

\subsubsection{Constructing constrains of uninterpreted 
predicate}
%
For a CDD system $S$, given a uninterpreted predicate $\cc{P} \in \predof{S}$, and a partial 
solution $\sigma$ that maps some uninterpreted predicate to its solution formula, 
,the formula $\ctrof{P,\sigma}$ is a compact representation of constrains of $\cc{P}$ based on parital
solution $\sigma$, defined as follows.
% for each relational predicate, define constraint:
If partial solution $\sigma$ does not contains $P$, 
the constraint of $P$ under $\sigma$ that denoted as $\ctrof{P,\sigma}$ 
is constructed as: for all clauses $C \in S$ in which $P$ is the head of $\cc{C}$
\textbf{(1)} each predicate $Q \in \bodyof{C}$ is used and 
\textbf{(2)} the constrained of the clause $\consof{C}$ is encoded.
%
I.e., $\ctrof{P,\sigma}$ is
\[
\biglor_{ (\cc{C}_i \in \cc{S}) \land (\headof{C_i} = \cc{P})}
\left( \consof{C_i} \land %
\bigland_{ \cc{Q} \in \bodyof{C_i}} %
b_Q \right)
\]
%
If partial solution $\sigma$ contains $P$, then $\ctrof{P,\sigma}$ is the solution formula $\sigma[P]$.

% introduce the isused predicates:
For each $\cc{P} \in \predof{S}$, there is a corresponding boolean variable
$b_P$.
%
The counterexample characterization of $P$ constrains that
if $P$ is used, which means $b_{P}$ is \cc{True}, then its constraint must hold.
%
I.e., $\vc{P,\sigma}$ is $\neg b_{P} \lor \ctrof{P,\sigma}$.

\begin{ex}
  \label{ex:ctr}
  In the CHC system $\mcchc$ (given in \autoref{sec:solve-ex}), 
  when $\mcsolve$ solves the uninterpreted predicate \cc{L9}, it 
  has a partial solution $\sigma$ that maps \cc{L6} to $n \ge 0$.
  %
  For predicate \cc{L6}, the the $\ctrof{L6,\sigma}$ is $n \ge 0$.
  %
  For the 
  uninterpreted \cc{L9}, the $\ctrof{L9,\sigma}$ is $(abs' = n \land b_{L6}) \lor (abs' = n \land b_{L8})$.
  %
  The $\vc{L9,\sigma}$ is $\neg b_{L9} \lor \ctrof{L9,\sigma}$.
\end{ex}

% pre-constraint:
\subsubsection{Constructing the pre-constraint of a relational predicate}
\label{sec:cons-pre}
%
For a CDD system $\cc{S}$, given a uninterpreted predicate $\cc{P} \in \predof{\cc{S}}$
, because $\solvecdd$ solve each uninterpreted predicate in a toplogical order, 
it has a partial solution $\sigma : \cc{Q} \to \formula$ for all uninterpreted predicates in
$\depsof{P}$.
%
the pre-constraint $\prectr(\cc{P},\sigma)$ is a formula that the conjunction of the constrains
of $\cc{P}$ and solutions of all dependencies of $P$ with their boolean varialbe.
%
In particular, $\prectr(\cc{P},\sigma)$ is:
\[\ctrof{P,\sigma} \land
\left(
\bigland_{\cc{Q} \in \depsof{P}}
\left(\neg b_Q \lor \sigma[ \cc{Q} ]  \right)
\right)
\]
%
\begin{ex}
  \label{ex:pre-ctr}
  In the CHC system $\mcchc$ (given in \autoref{sec:solve-ex}), 
  when $\mcsolve$ solves the uninterpreted predicate \cc{L9}, because it solves
  in topological order, partial solution $\sigma$ contains \cc{L6} and \cc{L8}.
  %
  $\sigma$ maps
  \cc{L6} to $n \ge 0$ and \cc{L8} to $n < 0$.
  %
  The pre-formula for \cc{L9} under $\sigma$ is $\ctrof{L9} \land (\neg b_{L6} \lor n \ge 0) \land (\neg b_{L8} \lor n < 0)$.
  %
  The forumula of $\ctrof{L9,\sigma}$ is given in~\autoref{ex:ctr}.
\end{ex}

\subsubsection{Constructing the post-constraint of a relational
  predicate}
\label{sec:cons-ctx}
%
For $\sigma$ as a partial solution of
$S$ and $P \in \predof{S}$, the post-constraint $\postctr(P, \sigma)$ is mutually
inconsistent with the solution of $P$ in each solution of
$S$ that extends $\sigma$.
% dependents of R:
In particular, let $D_0$ be the reflexive transitive dependents of $P$
in $S$,
% clause siblings of dependents:
let $D_1$ be the siblings in $S$ of
$(D_0 \union {P})$,
%
let $D_2$ be all transitive dependences of $D_1$, and
% collect all predicates that we define a constraint over:
let $D = D_0 \union D_1 \union D_2$.
% define instantiation of C
The post constraint for $P$ under $\sigma$ is the conjunction of
counterexample characterization of all predicates $Q \in D$ and the query clause.
%
I.e., $\postctr(R, \sigma) =\cc{query} \land (\bigland_{\cc{Q} \in D} \vc{Q,\sigma})$.
% example post constraint:
\begin{ex}
  \label{ex:ctx-ctr}
  Consider the case in which $\mcsolve$, given $\mcchc$, generates an
  interpretation for \cc{L9}, introduced in \autoref{ex:pre-ctr}.
  %
  The reflexive transitive dependents $D_0$ of \cc{L9} is \{\cc{main}\}.
  %
  The siblings set $D_1$ is \{\cc{db}\}.
  %
  The transitive dependences of $D_1$ is $\varnothing$. 
  %
  Then the $D$ set is \{\cc{main},\cc{db}\}
  %
  The \cc{query} is $b_{main} \land res'<0$.
  %
  The post constraint for $L9$ under $\sigma$ is $\cc{query} \land \vc{main,\sigma} \land \vc{db,\sigma}$.
  %
  The whole formula is given in\QZ{add reference}. 
\end{ex}
